---
title: "fasstr Under the Hood"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r options, include=FALSE}
#knitr::opts_chunk$set(eval = nzchar(Sys.getenv("hydat_eval")))
```

```{r, include=FALSE}
#library(fasstr)
```

### UNDER CONSTRUCTION

The Flow Analysis Summary Statistics Tool for R (`fasstr`) is a set of [R](http://www.r-project.org) functions to tidy, summarize, analyze, trend, and visualize streamflow data. This package summarizes continuous daily mean streamflow data into various daily, monthly, annual, and long-term statistics, completes annual trends and frequency analyses, in both table and plot formats.

This vignette guide contains a look at the internal steps and functions used complete the analyses in this package.

Cleaning Functions

- Argument checks
- flowdata_import - setting up data from data or station_number
- prepping - original columns, original grouping (will regroup at end), ungrouping
- formatting columns for analysis
- cleaning data with appropriate dates, water years and rolling day and cumulative flows for grouping and analyzing
- add columns or rows
- return original column names (and maybe remove STATION_NUMBER)
- return original grouping


Tables

- Argument checks
- flowdata_import - setting up data from data or station_number
- prepping - original columns and ungrouping
- formatting columns for analysis
- cleaning data with appropriate dates, water years and rolling day and cumulative flows for grouping and analyzing
- filtering for selected years and months
- summarizing (grouping by STATION_NUMBER)
- more filtering if necessary
- transposing
- warnings for missing values
- returning column names to original

Plots

- Argument checks
- flowdata_import - setting up data from data or station_number
- formatting columns for analysis
- calculate stats using functions
- setting some plotting objects (labels etc)
- create a tibble of plots
    - group by station_number
    - nest
    - new column using purrr to create plots
- grab column of plots into list with names
