---
title: "fasstr Users Guide"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
  
```{r options, include=FALSE}
knitr::opts_chunk$set(eval = nzchar(Sys.getenv("hydat_eval")))
```

fasstr, the Flow Analysis Summary Statistics Tool for R, is a set of [R](http://www.r-project.org) functions to tidy, summarize, analyze, trend, and visualize streamflow data. This package summarizes continuous daily mean streamflow data into various daily, monthly, annual, and long-term statistics, completes annual trends and frequency analyses, in both table and plot formats.

This vignette documents the usage of the many functions and arguments provided in fasstr. This vignette is a high-level adjunct to the details found in the various function documentations (see `help(package = "fasstr")` for documentations). You’ll learn how to install the package and a HYDAT database, input data into fasstr functions, add relevant columns and rows to daily data, screen data for outliers and missing dates, calculate and visualize various summary statisics, trend annual flows, and complete volume frequency analyses.

This guide contains the following sections to help understand the usage of the fasstr functions and arguments:

1. Getting Started
2. Flow Data Inputs
3. Function Outputs
4. Data Cleaning (`fill_*` and `add_*` functions)
5. Data Screening (`screen_*` functions)
6. Calculating Statistics (`calc_*` functions)
7. Analyses (`compute_*` functions)
8. Customizing Functions - Data filtering and options
9. Writing Tables and Plots (`write_*` functions)




## 1. Getting Started

### Installing and loading fasstr

To install the `fasstr` you will first need to install `devtools`, if not already done so, and then `fasstr` using the following code. It may take a few moments as there are several dependecy packages will also be installed, including [tidyhydat](https://cran.r-project.org/web/packages/tidyhydat/index.html) for downloading Water Survey of Canada hydrometric data, [zyp](https://cran.r-project.org/web/packages/zyp/index.html) for trending, [ggplot2](https://cran.r-project.org/web/packages/ggplot2/index.html) for creating plots, and [dplyr](https://cran.r-project.org/web/packages/dplyr/index.html) and [tidyr](https://cran.r-project.org/web/packages/tidyr/index.html) for various data wrangling and summarizing functions, amongst others. 

```{r, echo=TRUE, eval=FALSE}
install.packages("devtools")
devtools::install_github("bcgov/fasstr")
```

To call the `fasstr` functions you can either load the package using the `library()` function or access a specific function using a double-colon (e.g. `fasstr::calc_daily_stats()`). `fasstr` also exports the pipe, ` %>%`, so it can be used for tidy workflows. For this vignette, the `dplyr` package will also be installed.

```{r, echo=TRUE, message=FALSE, warning=FALSE}
library(fasstr)
library(dplyr)
```

### Downloading HYDAT

To use the `station_number` argument of the `fasstr` functions, you will need to download a HYDAT database to your computer using the following `tidyhydat` function. The function will save the database on your computer and know where to find it each time you open R or RStudio. Due to the size of the database, it will take several minutes to download.

```{r, echo=TRUE, eval=FALSE}
tidyhydat::download_hydat()
```

As HYDAT is updated frequently you may want to periodically update it yourself using the function above. You can check the local version using the following code:

```{r, echo=TRUE, eval=FALSE}
tidyhydat::hy_version()
```



## 2. Flow Data Inputs

All functions in `fasstr` require a daily mean streamflow dataset from one or more hydrometric stations. Long-term and continuous datasets are preferred for most analyses, but seasonal and partial data can be used. Data is provided to each function using one of the following arguments:

- `data`, as a data frame, or
- `station_number`, as a list of Water Survey of Canada HYDAT station numbers.

### 'data' (and dates, values, and groups)

Using the `data` option, a data frame of daily data containing columns of dates (YYYY-MM-DD in date format), values (mean daily discharge in cubic metres per second in numeric format), and, optionally, grouping identifiers (character string of station names or numbers) is called. By default the functions will look for columns identified as 'Date', 'Value', and 'STATION_NUMBER', respectively, to be compatible with the HYDAT default columns. However, columns of different names can be identified using the `dates`, `values`, `groups` column arguments (ex.  `values = Yield_mm`). The following is an example of an appropriate data frame with default column names (STATION_NUMBER not required):

```{r setup, include = FALSE}
data <- tidyhydat::hy_daily_flows("08NM116")
data <- data[,c(1,2,4)]
```

```{r flow_data, warning = FALSE, echo=FALSE}
head(data)
```

The following is an example fasstr calculation if your daily data data frame has the default columns names (no need to list them):

```{r, echo=TRUE, eval=FALSE}
calc_longterm_stats(data = flow_data)
```

The following is an example if your daily data data frame has the the columns names of "Stations", "Dates", and "Flows":

```{r, echo=TRUE, eval=FALSE}
calc_longterm_stats(data = flow_data,
                    dates = Dates,
                    values = Flows,
                    groups = Stations)
```

The `data` argument is listed first in the list of arguments for each function, so the data frame can be passed onto other functions using the pipe operator, `%>%`, in a tidy workflow.

### 'station_number'

Alternatively, you can directly extract flow data directly from a HYDAT database by listing station numbers in the `station_number` argument while leaving the data arguments blank. The following is an example of listing stations:

```{r, echo=TRUE, eval=FALSE}
calc_longterm_stats(station_number = "08NM116")
calc_longterm_stats(station_number = c("08NM116", "08NM242"))
```

This package allows for multiple stations (or other groupings) to be analyzed in many of the functions; provided they are identified using the `groups` column argument (defaults to STATION_NUMBER). If grouping column doesn't exist or is improperly named, then all values listed in the `values` column will be summarized.


## 3. Function Types and Outputs

fasstr provides various functions to help in streamflow analyses. They can be generally categorized into the following groups (with more details in the sections below):

- data cleaning (to prepare data for analyses; `add_*` and `fill_*` functions),
- data screening (to look for outliers and missing data; `screen_*` functions),
- calculating summary statistics (longterm, annual, monthly and daily statistics; `calc_*`functions),
- computing analyses (volume frequency analyses and trending; `compute_*` functions),
- visualizing (plotting the various statistics; `plot_*` functions), and
- writing (to save your data and plots; `write_*` functions)

### Tibbles

Functions that produce tables create them as tibble data frames, which are a more modern take on data frames. More information on tibbles can be found [here](https://cran.r-project.org/web/packages/tibble/vignettes/tibble.html). To facilitate the writing of the fasstr tibbles to a directory as .csv, .xls, or .xlsx files with some functionality of rounding digits, the `write_results()` function can be used (see section 9 for more information)..

### ggplots

Functions that produce plots create them as lists of ggplot2 objects. The use of ggplots allows for further customization of plots for the user (axis titles, colours, etc.). All plotting functions produce lists to be consistent with table naming conventions of fasstr, allow multiple plots to be created with one function, and to easily allow the saving of multiple plots to a directory. To assist with the saving of lists of plots, a provided function called `write_plots()` will directly save the list of plots within a directory or single PDF document, with the fasstr plot objects names (see section 9 for more information).  Individual plots can be subsetted from their lists using either the dollar sign, \$ (e.g. `one_plot <- plots$plotname`), or double square brackets, [ ] (e.g. `one_plot <- plots[[plotname]]` or `one_plot <- plots[[1]]`).

Some functions produce both tibbles and plots as lists, and can be subsequently subsetted as desired.


## 4. Data Cleaning Functions

There are several functions that are used to prepare your flow dataset for your own analysis. These functions start with `add_` or `fill_` and add columns or rows, respectively, to your flow data frame. These functions include:

- `fill_missing_dates()` - fills in missing dates or dates with no flow values with NA
- `add_date_variables()` - add year, month, and day of year variables (and water years if selected)
- `add_seasons()` - add two and four seasons identifier columns
- `add_rolling_means()` - add rolling n-day averages
- `add_basin_area()` - add a basin area column to daily flows
- `add_daily_volume()` - add daily volumetric flows (in cubic metres)
- `add_daily_yield()` - add daily volumetric runoff yields (in millimetres)
- `add_cumulative_volume()` - add daily cumulative volumetric flows on an annual basis (in cubic metres)
- `add_cumulative_yield()` - add daily cumulative runoff yield flows on an annual basis (in millimetres)

The functions are set up to easily incorporate the use of the pipe operator, like the following:

```{r exampletidy, warning=FALSE}
fill_missing_dates(station_number = "08HA011") %>% 
  dplyr::select(Date, Value) %>%  # keeps just the Date and Value columns
  add_date_variables() %>% 
  add_rolling_means(roll_days = 7)
```

### Filling missing dates

To ensure that analyses do not skip over dates, the `fill_missing_dates()` function looks for gaps in dates and adds the dates and fills in the flow values with NA. It does not do any gap filling (linear or correlations, for example), it just gives those values NA. It also fills dates to create complete start and end years. For example, if data starts in April, all flow values starting from January will be filled with NA. The timing of the year depends on the `water_year` and `water_year_start` arguments. When `water_year = FALSE` (default), it will fill to complete calendar years (Jan-Dec). If `water_year = TRUE` then it will fill to complete water years, defaulting to Oct-Sep, or other year periods with a different starting month with `water_year_start`.

Run and compare the following lines to see how missing dates are filled:

```{r, warning=FALSE, eval=FALSE}
# Very gappy (early years):
tidyhydat::hy_daily_flows(station_number = "08NM116")

# Gap filled with NA's
tidyhydat::hy_daily_flows(station_number = "08NM116") %>% 
  fill_missing_dates()
```

It is ideal to fill missing dates before using other `add_*` functions so dates added are not missing the other new values.

### Adding date variables and seasons

The `add_date_variables()` function added useful dates columns for summarizing data. The function defaults include 'Year' (calendar year), 'Month' (numeric), 'MonthName' (month abbreviation; e.g. Jan), and 'DayofYear' (the day of year from 1-365). If `water_year = TRUE` for this function then it will also add 'WaterYear' and 'WaterDayofYear' (starting on the first day of the water year) columns in addition. Water year will start depending on the `water_year_start` argument, that defaults to "9" for October.

Run and compare the following lines to see how the date columns are added:

```{r, eval=FALSE}
# Just calendar year info
add_date_variables(station_number = "08NM116")

# If water years are require (default starting October)
add_date_variables(station_number = "08NM116", water_year = TRUE)

# If water years are require starting August (use month number)
add_date_variables(station_number = "08NM116", water_year = TRUE, water_year_start = 8)
```

The `add_seasons()` functions added two columns of season identifiers: Seasons2 (two 6-month seasons) and Seasons4 (four 3-month seasons). The start of the two and four seasons coincides with the start month of each year ('Jan-Jun' and 'Jan-Mar' for the first two and four calendar year seasons, respectively, or 'Oct-Mar' and 'Apr-Sep' for the first two and four water year seasons starting in Oct, respectively). Each season is designated by the calendar or water year in which it occurs.

Run and compare the following lines to see how seasons columns are added:

```{r, eval=FALSE}
#  Seasons starting January
add_seasons(station_number = "08NM116")

#  Seasons starting October
add_seasons(station_number = "08NM116", water_year = TRUE)

#  Seasons starting December
add_seasons(station_number = "08NM116", water_year = TRUE, water_year_start = 12)
```


### Adding rolling means

Adding rolling means, or running means or averages of daily data can be done unsing the `add_rolling_means()` functions. Based on the selected "n" rolling days using the `roll_days` argument, a column for each "n" will be added. One rolling mean column can be added by listing one number (e.g. `roll_days = 7`) or multiple columns can be added by listing each one (e.g. `roll_days = c(3,7,30)`).

Where the alignment of the rolling mean is compared to the date is important to know when analyzing data.  The alignment, using the `roll_align` argument, determine the date at which the rolling means occur. If `roll_align = "right"`, the default, then the date will have the mean of that date's flow value and the previous n-1 days. If `roll_align = "left"`, then the date will have the mean of that date's flow value and the next n-1 days. If `roll_align = "center"`, then the date will be in the center of the rolling values. For odd roll_days with center alignment, the date will have the mean of that date's flow value and half of n-1 days before and half of n-1 days after. For even roll_days with center alignment, the date will have the mean of that date's flow and half of n days after, and the remaining before ((n/2)-1 days before the date).

- `roll_align = "right"` - 
- `roll_align = "left"` - 
- `roll_align = "center"` 
    - odd numbered `roll_days` - 
    - even numbered `roll_days` - 

For example, the 3 days with "left" show the date is on the "left" or first of all 5 flow values. The 4 days with "center" show the date is in the centre of all 5 flow values (if an even number, then half the flows are after the date, and the other half are the date and the previous flow values). The 7 days with "right" show the date is on the "right" or last of all 7 flow values.

```{r, warning=FALSE, eval=TRUE}
add_rolling_means(station_number = "08HA011", roll_days = 3, roll_align = "left") %>% 
  add_rolling_means(roll_days = 4, roll_align = "center") %>% 
  add_rolling_means(roll_days = 7, roll_align = "right")
```

```{r, warning=FALSE, eval=TRUE, echo=FALSE}
head(add_rolling_means(station_number = "08HA011", roll_days = 5, roll_align = "left") %>% 
  dplyr::rename("Q5Day_left" = Q5Day) %>% 
  add_rolling_means(roll_days = 5, roll_align = "center") %>% 
  dplyr::rename("Q5Day_center" = Q5Day) %>% 
  add_rolling_means(roll_days = 5, roll_align = "right") %>% 
  dplyr::rename("Q5Day_right" = Q5Day) %>% 
  dplyr::select(-STATION_NUMBER, -Parameter, -Symbol))

```

```{r, warning=FALSE, eval=TRUE, echo=FALSE}
head(add_rolling_means(station_number = "08HA011", roll_days = 6, roll_align = "left") %>% 
  dplyr::rename("Q6Day_left" = Q6Day) %>% 
  add_rolling_means(roll_days = 6, roll_align = "center") %>% 
  dplyr::rename("Q6Day_center" = Q6Day) %>% 
  add_rolling_means(roll_days = 6, roll_align = "right") %>% 
  dplyr::rename("Q6Day_right" = Q6Day) %>% 
  dplyr::select(-STATION_NUMBER, -Parameter, -Symbol))

```

### Adding basin areas


Run and compare the following lines to see how basin area columns are added:

```{r, eval=FALSE}
# If you have a column with HYDAT station number (with station_number or data/groups arguments) then it will automatically look up the number
add_basin_area(station_number = "08NM116")

# If no basin area provided or you want to use a different one provided, sue the basin_area argument
add_basin_area(station_number = "08NM116", basin_area = 800)

# If you have multiple stations you need to apply basin_areas to, use the basin_area argument as a list and name each station
add_basin_area(station_number = c("08NM116","08NM242"), basin_area = c("08NM116" = 800, "08NM242" = 4))
```

### Adding daily volumetric or yield flows

Add columns of total dishcharge converted from daily mean into volumetric flows, named "Volume_m3" in cubic metres per second, or runoff yield, named "Yield_mm" in millimetres. Volumetric gives the total volume per day, and the yield gives the total water depth, provided an upstream drainage basin area is provided. Basin area can be prpvoded using the `basin_area` argument, or if there is a `groups` column of HYDAT station numbers in your data then it will automatically be etracted from HYDAT, if available. (see section XXXX for more information on basin areas).

```{r, eval=FALSE}
# Add a column of converted disharge (m3/s) into volume (m3)
add_daily_volume(station_number = "08NM116")    

# Add a column of converted disharge (m3/s) into yield (mm), with HYDAT station number groups
add_daily_yield(station_number = "08NM116")   

# Add a column of converted disharge (m3/s) into yield (mm), with setting the basin area
add_daily_yield(station_number = "08NM116", basin_area = 800)   
```


### Adding annual cumualtive daily volumetric or yield flows

These functions create a rolling cumulative of daily total flows on an annual basis, as volumetric flows, named "Cumul_Volume_m3" in cubic metres per second, or runoff yield flows, named "Cumul_Yield_mm" in millimetres. A total flow for a given a day is the sum of all previous days and that day, within a given year (Jan 15 cumulative flow value is the sum of all total flows from Jan 1-15). It restarts for each year (selected calendar or water year) and no values are calculated if there is missing data for a given year.

```{r, eval=FALSE}
# Add a column of converted disharge (m3/s) into cumulative volumes (m3)
add_cumulative_volume(station_number = "08NM116") 

# Add a column of converted disharge (m3/s) into cumulative yield (mm), with HYDAT station number groups
add_cumulative_yield(station_number = "08NM116")   

# Add a column of converted disharge (m3/s) into cumulative yield (mm), with setting the basin area
add_cumulative_yield(station_number = "08NM116", basin_area = 800)  
```


## 5. Data Screening

If you are looking at some data for the first time, it may be useful to explore the data quality and availability. The following fuctions will help to explore the data:

- `screen_flow_data()` - calculate annual summary and missing data statistics
- `plot_data_screening()` - plot annual summary statistics for data screening
- `plot_missing_dates()` - plot annual and monthly missing dates

Screening does this. To view that data, you can plot the data screening and plot the missing dates.


## 6. Calculating Statistics

The majority of the `fasstr` functions produce statistics over a certain time period, either long-term, annually, monthly, or daily. These statistics are produced using the `calc_*` functions and can be visualized using their corresponding `plot_*` functions. The following sections are an overview of the different statistic functions.

### Basic Summary Statistics

Basic summary statistics that contain means, medians, maximums, minimums, and percentiles:

- `calc_longterm_stats()` - calculate the long-term and long-term monthly summary statistics
- `calc_annual_stats()` - calculate annual summary statistics
- `calc_monthly_stats()` - calculate monthly summary statistics
- `calc_daily_stats()` - calculate daily summary statistics

and their corresponding plotting functions:

- `plot_longterm_stats()` - plot the long-term and long-term monthly summary statistics
- `plot_annual_stats()` - plot annual summary statistics
- `plot_monthly_stats()` - plot monthly summary statistics
- `plot_daily_stats()` - plot daily summary statistics

Mean, med, max min are standards, and percentiles can be selected.

Plotting main statistics. Long-term, monthly, and daily come pre-set with percentiles and cannot be adjusted.  If listed `values` column is either 'Volume_m3' or 'Yield_mm' (from their respective `add_*` functions) then the discharge axis title will adjust accordingly.


Can also plot flow duration curves using the function:

- `plot_flow_duration()` - plot flow duration curves

### Cumulative Flow Statistics

These provide the basic summary statistics of annual cumulative flows, volumetric or runoff yield for each period. 

- `calc_annual_cumulative_stats()` - calculate annual (and seasonal) cumulative flows
- `calc_monthly_cumulative_stats()` - calculate cumulative monthly flow statistics
- `calc_daily_cumulative_stats()` - calculate cumulative daily flow statistics

and their corresponding plotting functions:

- `plot_annual_cumulative_stats()` - plot annual and seasonal total flows
- `plot_monthly_cumulative_stats()` - plot cumulative monthly flow statistics
- `plot_daily_cumulative_stats()` - plot cumulative daily flow statistics

For example, the annual provide the total annual and seasonal (optional) flows, and the monthly and daily period cumulate for each month or day until the end of the year. These functions default to volumetric flows, but using the `use_yield` and `basin_area` arguments, the runoff yield can be calculated for each period. If there is a `groups` column of HYDAT station numbers, then the function will automatically pull the basin area out of HYDAT if available; otherwise a basin area will be required. Only calculated using years of complete data.

Plotting statistics. Monthly and daily come pre-set with percentiles and cannot be adjusted.


### Various Annual Statistics

These are other annual statistic functions that produce....:

- `calc_annual_flow_timing()` - calculate annual flow timing
- `calc_annual_lowflows()` - calculate annual lowflows
- `calc_annual_outside_normal()` - calculate annual days above and below normal
- `calc_all_annual_stats()` - calculate all fasstr annual statistics

and their corresponding plotting functions:

- `plot_annual_flow_timing()` - plot annual flow timing
- `plot_annual_lowflows()` - plot annual lowflows
- `plot_annual_outside_normal()` - plot annual days above and below normal

The `calc_annual_flow_timing()` calculates the day of year when a portion of an total annual volumetric flow has occurred. Using the `percent_total` argument, one or multiple portions of annual flow can be calculated. Using 50 as the percent is similar to the centre of volume (see link?). The day of year and date will be produced.

The `calc_annual_lowflows()` calculates the annual minimum values, day of years, and dates of specified rolling mean days. 

The `calc_annual_outside_normal()` calculates the number of days per year that are above and below "normal", "normal" typically defined as 25th and 75th percentiles. The normal limits can be determined using the `normal_percentiles` argument, listing the lower and upper normal ranges, respectively (e.g. c(25, 75)). The function calculates the lower and upper percentiles for each day of the year over all years, and sums all days that are above or below the daily normal ranges for a given year.

The `calc_all_annual_stats()` calculates all statistics that have a single annual value. This includes all the `calc_annual_*` and the `calc_monthly_statistics()` functions. Several argumements provided for customization of the statistics. There is no corresponding plotting function for this calculation function.


### Single-value Statistics

The functions summarize the data over the entire record:

- `calc_lt_mad()` - calculate the long-term mean annual discharge
- `calc_lt_percentile()` - calculate the long-term percentiles
- `calc_flow_percentile()` - calculate the percentile rank of a flow value

`calc_lt_mad()` calculates the mean of all the daily flows. It can also be known as the long-term mean annual discharge, MAD.

`calc_lt_percentile()` calculates the selected long-term percentiles of all the daily flow values.

`calc_flow_percentile()` calculates the percentile rank of a specified flow value, as `flow_value`. It compares all daily flow values determines where the specified flow values ranks amongst all flows.


## 7. Analyses

Trending:

- `compute_annual_trends()` - calculate prewhitened nonlinear annual trends on streamflow data
- `plot_annual_trends()` - plot prewhitened nonlinear annual trends on streamflow data

Volume Frequency: 

- `compute_annual_frequencies()` - perform an annual low or high-flow frequency analysis
- `compute_frequency_stat()` - calculate an annual frequency analysis statistic
- `compute_hydat_peak_frequencies()` - perform a flow frequency analysis on annual statistics


## 8. Customizing Functions - Data filtering and options

One of the strengths of fasstr is the ability to customize how data is filtered and analyzed within functions, thereby minimizing the lines of code in your scripting.
Described here are some of the options available to the functions of this package. Not all functions have all these options, see the documentation for each function to see which argument options are available.


### Missing Dates

Dealing with data gaps is important ...

Most functions will automatically (`ignore_missing = FALSE`) not calculate a statistic for that period if there is missing data (base `na.rm = FALSE`) and will produce an NA value or will not plot. For example, if there at least one missing day for a given year, an annual statistic will not be calcuated for that year. A warning message will appear in the console indicating as such. See the following code for an example with missing dates:

```{r, eval=FALSE}
calc_annual_stats(station_number = "08NM116")
```

If you want to calculate the statistics regardless of missing dates, use the `ignore_missing = TRUE` argument, as such:

```{r, eval=FALSE}
calc_annual_stats(station_number = "08NM116", 
                  ignore_missing = TRUE)
```


### Dates Filtering

#### Water Year and Start Month

May want a different time period

To have years starting in October
```{r, eval=FALSE}
calc_annual_stats(station_number = "08NM116", 
                  start_year = 1980, end_year = 2010,
                  water_year = TRUE)
```

To have years starting in August
```{r, eval=FALSE}
calc_annual_stats(station_number = "08NM116", 
                  start_year = 1980, end_year = 2010,
                  water_year = TRUE,
                  water_year_start = 8)
```

#### Start and End Years

If you have missing data in certain years or you have a specific period to analyze, you can filter using several options:

Start and end years
```{r, eval=FALSE}
calc_annual_stats(station_number = "08NM116", 
                  start_year = 1980, 
                  end_year = 2010)
```

#### Exclude Years

# Removing certain years (outliers, bad data, etc) using exclude_years
```{r, eval=FALSE}
calc_annual_stats(station_number = "08NM116", 
                  start_year = 1980, 
                  end_year = 2010,
                  exclude_years = 1982)
```

```{r, eval=FALSE}
calc_annual_stats(station_number = "08NM116", 
                  start_year = 1980, 
                  end_year = 2010,
                  exclude_years = c(1982:1984))
```

#### Complete Years

If your data has missing dates, but you would like to use only those years with complete data, some functions utilize the `complete_years` argument where the data will automatically be filtered for years with complete data and statistics will be calculated. The following code will only calculte the longterm statistics with years of complete data.

```{r, eval=FALSE}
calc_longterm_stats(station_number = "08NM116", complete_years = TRUE)
```

Some functions require years with complete data, so years with gaps will be ignored:
```{r, eval=FALSE}
calc_annual_flow_timing(station_number = "08NM116")
```

#### Months

# Some functions allow you use data from selected months
# This example shows the annual stats for just June-August (Default 1:12)
```{r, eval=FALSE}
calc_annual_stats(station_number = "08NM116", 
                  start_year = 1980, end_year = 2010,
                  months = 6:8)
```


### Rolling Days


# Some functions (mostly calc_***_stats() functions) allow you to analyze rolling mean days instead of 
# the daily mean (roll_days default is 1)

```{r, eval=FALSE}
# This calculates the annual 7-day rolling mean statistics
calc_annual_stats(station_number = "08NM116", 
                  start_year = 1980, end_year = 2010,
                  roll_days = 7)
plot_annual_stats(station_number = "08NM116", 
                  start_year = 1980, end_year = 2010,
                  roll_days = 7)
```

```{r, eval=FALSE}

# This calculates the annual 7-day rolling mean statistics with align left (see add_rolling_days() above)
calc_annual_stats(station_number = "08NM116", 
                  start_year = 1980, end_year = 2010,
                  roll_days = 7,
                  roll_align = "left")
plot_annual_stats(station_number = "08NM116", 
                  start_year = 1980, end_year = 2010,
                  roll_days = 7,
                  roll_align = "left")
```


Roll days

Roll align

Some just one.


### Percentiles and Other Statistics

Mean/Median/Max/min

Percentiles

Others

Functions come with their default settings for what is calculated and they can be changed with argument setting.
Look at the help files to see default and arguments

calc_annual_stats defauls to Mean, Meidan, Maximum, and Minimum stats with 10th and 90th percentiles
you can change just the percentiles using the argument
```{r, eval=FALSE}
calc_annual_stats(station_number = "08NM116", 
start_year = 1980, end_year = 2010,
percentiles = c(1:5,20,30))
plot_annual_stats(station_number = "08NM116", 
start_year = 1980, end_year = 2010,
percentiles = c(60,70,80))
```

For annual flow timing, you can change the percent of total annual flow you want to know the date of:
Default 25, 33.3, 50, and 75 percent of total flow
```{r, eval=FALSE}
calc_annual_flow_timing(station_number = "08NM116", 
start_year = 1980, end_year = 2010)
# Just timing of half flow
calc_annual_flow_timing(station_number = "08NM116", 
start_year = 1980, end_year = 2010, 
percent_total = 50)
```


#### Table Options

Transpose


#### Plot Options

Log discharge
Normal scale:
```{r, eval=FALSE}
plot_annual_stats(station_number = "08NM116", 
start_year = 1980, end_year = 2010)
```

Normal scale:
```{r, eval=FALSE}
plot_annual_stats(station_number = "08NM116", 
start_year = 1980, end_year = 2010,
log_discharge = TRUE)
```

Titles

Yiled or volume as value

Add year

Some plotting functions provide the ability to plot indiviual years on plots where stats from all years are summarize
(ex. plot_daily_stats, plot_daily_cumulative stats, plot_monthly_cumulative_stats)
```{r, eval=FALSE}
plot_daily_stats(station_number = "08NM116", complete_years = TRUE,
include_year = 2010)
```



## 9. Writing Tables and Plots

Writing analyses:

- `write_full_analysis()` - write a suite of tables and plots of various statistics

Writing data and plots:

- `write_flow_data()` - write a streamflow dataset as a .xlsx, .xls, or .csv file
- `write_plots()` - write plots from a list into a directory or PDF document
- `write_results()` - write a data frame as a .xlsx, .xls, or .csv file



### WHERE TO PUT ME?!

Produce plots:

- `plot_flow_data()` - plot daily mean streamflow





### Summary statistics example: long-term statistics

To determine the summary statistics of an entire dataset and by month (mean, median, maximum, minimum, and some percentiles) you can use the `calc_longterm_stats()` function. If the 'Mission Creek near East Kelowna' hydrometric station is of interest you can list the station number in the `HYDAT` argument to obtain the data (if `tidyhydat` and HYDAT are installed). 

```{r example1, warning=FALSE, eval=FALSE}
calc_longterm_stats(station_number = "08NM116")
```

### Plotting example 1: daily summary statistics

To visualize the daily streamflow patterns on an annual basis, the `plot_daily_stats()` function will plot out various summary statistics for each day of the year. Data can also be filtered for certain years of interest (a 1981-2010 normals period for this example) using the `start_year` and `end_year` arguments. Multiple plots are produced with this function, so this example plots just the summary statistics (`[1]`).

```{r plot1, warning=FALSE, eval = FALSE, fig.height = 4, fig.width = 10}
plot_daily_stats(station_number = "08NM116",
                 start_year = 1981,
                 end_year = 2010,
                 log_discharge = TRUE)[1]
```

### Plotting example 2: flow duration curves

Flow duration curves can be produced using the `plot_flow_duration()` function.

```{r plot2, warning=FALSE, eval = FALSE, fig.height = 4, fig.width = 7}
plot_flow_duration(station_number = "08NM116",
                   start_year = 1981,
                   end_year = 2010)
```

### Analysis example: low-flow frequency analysis

This package also provides a function, `compute_frequency_analysis()`, to complete frequency analyses (using the same methods as [HEC-SSP](http://www.hec.usace.army.mil/software/hec-ssp/)). The default fitting distribution is 'log-Pearson Type III', but the 'weibull' distribution can also be used. Other default plotting and fitting methods are described in the function documentation. For this example, the 7-day low-flow (low-flow is default) quantiles are calculated for the Mission Creek hydrometric station using the 'log-Pearson Type III' distribution. With this, several low-flow indicators can be determined (i.e. 7Q5, 7Q10).

```{r example2, eval= FALSE, warning=FALSE}
compute_frequency_analysis(station_number = "08NM116",
                           start_year = 1981,
                           end_year = 2010,
                           rolling_days=7)[5]
```

