---
title: "Using USGS dataRetrieval with fasstr"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using USGS dataRetrieval with fasstr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r options, include=FALSE}
knitr::opts_chunk$set(eval = nzchar(Sys.getenv("hydat_eval")),
                      warning = FALSE, 
                      message = FALSE)
```

`fasstr`, the Flow Analysis Summary Statistics Tool for R, is a set of [R](https://www.r-project.org/) functions to tidy, summarize, analyze, trend, and visualize streamflow data. This package summarizes continuous daily mean streamflow data into various daily, monthly, annual, and long-term statistics, completes trending and frequency analyses.

This vignette explores how to obtain and integrate U.S. Geological Survey (USGS) streamflow data (available through the [National Water Information System (NWIS)](https://waterdata.usgs.gov/nwis)) using the [`dataRetrieval` R package](https://cran.r-project.org/package=dataRetrieval) with the various `fasstr` functions.

This vignette will use the following packages:

```{r}
library(dataRetrieval)
library(tidyhydat)
library(fasstr)
library(dplyr)
```

### dataRetrieval

`dataRetrieval` an R package with a collection of functions to retrieve hydrologic and water quality data hosted on USGS and U.S. Environmental Protection Agency (EPA) web services. USGS hydrologic data are pulled from <https://waterservices.usgs.gov/> and <https://waterdata.usgs.gov/nwis>. Water quality data are obtained from the Water Quality Portal <https://www.waterqualitydata.us/>. More information on using `dataRetrieval` can be found in the [vignette](https://cran.r-project.org/web/packages/dataRetrieval/vignettes/dataRetrieval.html).

This package is another useful tool for hydrologists, especially for those looking to use USGS streamflow hydrometric station data with `fasstr` or other streamflow packages and analyses. You can search for stations on the NWIS website <https://waterdata.usgs.gov/nwis/rt>. Unlike the `tidyhydat` package to obtain  Environment and Climate Change Canada's (ECCC) HYDAT historical discharge data, `dataRetrieval` pulls all data from the NWIS web services and does not require downloading and storing a database.

As USGS stations collect more than just streamflow, a parameter code is required for many of the functions to choose the appropriate data type. The parameter code for daily mean discharge data is "00060". The units of this discharge parameter are in US customary cubic feet per second (cfs). While using US customary units is not an issue for many `fasstr` functions, as some plots titles or column headers can simply be renamed, there are some functions that do metric unit conversions so some adaptations will be required (see sections below).

Here is a quick workflow demo showing how to get hydrometric station information, and daily and annual peak data using `dataRetrieval`, adapted from the [vignette](https://cran.r-project.org/web/packages/dataRetrieval/vignettes/dataRetrieval.html):

```{r, eval=FALSE}
# Hoko River Near Sekiu, WA (12043300)
site_number <- "12043300" 

# Get site information (ie. name, coordinates, drainage basin area, etc.)
site_info <- readNWISsite(site_number)

# Get daily discharge data
# Requires parameter code for discharge (00060), in cubic feet per second (cfs)
usgs_data <- readNWISdv(siteNumbers = site_number,
                        parameterCd = "00060", 
                        startDate = "1980-01-01",
                        endDate = "2019-12-31")

# Get annual instantaneous peak discharge data (based on water years (Oct-Sep))
peak_data <- readNWISpeak(siteNumbers = site_number)

```


### Converting to metric and tidyhydat/fasstr structure

For metric discharge analyses using USGS data, discharge values can easily be converted to cms from cfs, and for ease with use of fasstr functions, column names be renamed to match those with the default HYDAT data structure (ie. 'STATION_NUMBER', 'Date', and 'Value'). 

The default dataRetrieval structure is as follows:

```{r, eval=TRUE}
usgs_data <- readNWISdv(siteNumbers = "12043300", parameterCd = "00060")
head(usgs_data)
```

If you desire to keep the USGS data column names, just the convert the discharge units and send to fasstr functions using the 'values' argument. Date column names are the same in both NWIS and HYDAT data. Both base R and pipeline examples are as follow:

```{r, eval=FALSE}
# Get cfs data
usgs_data <- readNWISdv(siteNumbers = "12043300",
                        parameterCd = "00060")

# Convert cfs to cms
usgs_data$X_00060_00003 <- usgs_data$X_00060_00003 * 0.028317

# Set `values` and 'groups' (if multiple stations) to the appropriate column names
plot_flow_data(data = usgs_data,
               values = X_00060_00003,
               groups = site_no)

# Same process within a pipeline
readNWISdv(siteNumbers = "12043300", 
           parameterCd = "00060") %>% 
  mutate(Value = Value * 0.028317) %>% 
  plot_flow_data(values = X_00060_00003,
                 groups = site_no)
```

The USGS data can also be easily converted into HYDAT formatting to send to fasstr functions with ease. Renaming  'site_no' to 'STATION_NUMBER' is not necessary if one station is used, but is useful if data contains multiple stations. Using base R:

```{r, eval=FALSE}
# Get the data
usgs_data <- readNWISdv(siteNumbers = "12043300", parameterCd = "00060")

# Rename columns
names(usgs_data)[names(usgs_data) == 'site_no'] <- 'STATION_NUMBER'
names(usgs_data)[names(usgs_data) == 'X_00060_00003'] <- 'Value'

# Convert cfs to cms
usgs_data$Value <- usgs_data$Value * 0.028317

# Use a fasstr function
calc_annual_peaks(usgs_data)
```

Alternatively, renaming columns and unit conversions using dplyr and the pipeline:

```{r, eval=FALSE}
readNWISdv(siteNumbers = "12043300", 
           parameterCd = "00060") %>% 
  rename(STATION_NUMBER = site_no, 
         Value = X_00060_00003) %>% 
  mutate(Value = Value * 0.028317) %>% 
  calc_annual_peaks()
```

Functions can also be created if conversions will be frequently occur within workflows. The first example is a function to convert data already downloaded using dataRetrieval's `readNWISdv()` function:

```{r, eval=FALSE}
# Create function to convert to metric and HYDAT structure
usgs_to_hydat <- function(data){
  names(data)[names(data) == 'site_no'] <- 'STATION_NUMBER'
  names(data)[names(data) == 'X_00060_00003'] <- 'Value'
  data$Value <- data$Value * 0.028317
  return(data)
}

# Get data, convert, and send to fasstr function (base and pipeline)
usgs_data <- readNWISdv("12043300")
usgs_data <- usgs_to_hydat(usgs_data)
calc_annual_stats(usgs_data)

readNWISdv("12043300") %>% 
  usgs_to_hydat() %>% 
  calc_annual_stats()
```

This second function integrates `readNWISdv()` with the first function that will both download and convert the data for fasstr functions by supplying USGS site numbers:

```{r, eval=FALSE}
# Create function to download USGS data and convert to metric and HYDAT structure
readNWISdv_hydat <- function(siteNumbers){
  data <- readNWISdv(siteNumbers = siteNumbers, parameterCd = "00060")
  names(data)[names(data) == 'site_no'] <- 'STATION_NUMBER'
  names(data)[names(data) == 'X_00060_00003'] <- 'Value'
  data$Value <- data$Value * 0.028317
  return(data)
}

# Get data and send to fasstr function (base and pipeline)
usgs_data <- readNWISdv_hydat("12043300")
calc_annual_stats(usgs_data)

readNWISdv_hydat("12043300") %>% 
  calc_annual_stats()
```

Some functions will also require a drainage basin area to calculate discharge as area-based runoff yield (in depth units) statistics. The USGS upstream drainage basin area can be found in the site information using the dataRetrieval `readNWISsite()` function and converted from square miles to square kilometres:

```{r, eval=FALSE}
# Get station information
site_info <- readNWISsite("12043300")

# Get the drainage area and convert to sq. km
basin_area_USGS <- site_info$drain_area_va * 2.58999

# Get data and apply to fasstr function
readNWISdv(siteNumbers = "12043300", 
           parameterCd = "00060") %>% 
  rename(Value = X_00060_00003) %>% 
  mutate(Value = Value * 0.028317) %>% 
  plot_daily_cumulative_stats(basin_area = basin_area_USGS, 
                              use_yield = TRUE)
```


### Working in US customary units

in usgs data and water years (water year designated by the calendar year it ends in, ie. 2010 for water year 2009-2010)

using default names. good for one off, but applying to mulitple fucntions easier to change to Value and STATION)NUMBER

```{r, eval=FALSE}
# Download and format USGS data
usgs_data <- readNWISdv(siteNumbers = "12043300", 
                        parameterCd = "00060")

calc_longterm_mean(data = usgs_data, 
                   values = X_00060_00003,
                   water_year_start = 10, 
                   complete_years = TRUE)
```


```{r, eval=FALSE}
# Download and format USGS data
usgs_data <- readNWISdv(siteNumbers = "12043300", 
                        parameterCd = "00060") %>% 
  rename(Value = X_00060_00003)

# Fill in some data gaps and add some columns
usgs_data <- usgs_data %>% 
  fill_missing_dates(water_year_start = 10) %>% 
  add_date_variables(water_year_start = 10) %>% 
  add_rolling_means(roll_days = 7)

# Send to various fasstr functions
screen_flow_data(usgs_data, 
                 water_year_start = 10)

calc_longterm_daily_stats(usgs_data, 
                          water_year_start = 10,
                          complete_years = TRUE)

calc_longterm_mean(usgs_data, 
                   water_year_start = 10, 
                   complete_years = TRUE)

calc_annual_lowflows(usgs_data, 
                     water_year_start = 10)
```


Using multiple stations
```{r, eval=FALSE}
# Download data from two stations
usgs_data <- readNWISdv(siteNumbers = c("12043300","12048000"), 
                        parameterCd = "00060") %>% 
  rename(Value = X_00060_00003,
         STATION_NUMBER = site_no)

# Send to fasstr
calc_longterm_daily_stats(usgs_data, complete_years = TRUE)
```



-   use usgs columns names example
-   rename to value for ease
-   Wrangling: convert dataRetrieval to tidy fasstr

#### Examples

-   basic, no basin_area
-   Simple calcs and plots (no conversions), screen, plot, plot_means, daily, low flow frequency, flow duration, trending?, custom peak frequency

#### Frequency Examples

#### Basin Area Examples

-   area-based runoff/yield and cumulative examples
-   basin area tricks

USE WASHINGTON EXAMPLES - COAST AND INTERIOR

12048000 DUNGENESS RIVER NEAR SEQUIM, WA medium coastal, very mixed regime 12045500 ELWHA RIVER AT MCDONALD BR NEAR PORT ANGELES, WA large coastal, mixed 12447390 ANDREWS CREEK NEAR MAZAMA, WA small nival 12404500 KETTLE RIVER NEAR LAURIER, WA large nival 12442500 SIMILKAMEEN RIVER NEAR NIGHTHAWK, WA nival 12043300 HOKO RIVER NEAR SEKIU, WA small coastal, pluvial 12212050 FISHTRAP CREEK AT FRONT STREET AT LYNDEN, WA pluvial - Abbotsford
