---
title: "Using USGS dataRetrieval with fasstr"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using USGS dataRetrieval with fasstr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r options, include=FALSE}
knitr::opts_chunk$set(eval = nzchar(Sys.getenv("hydat_eval")),
                      warning = FALSE, 
                      message = FALSE)
```

`fasstr`, the Flow Analysis Summary Statistics Tool for R, is a set of [R](https://www.r-project.org/) functions to tidy, summarize, analyze, trend, and visualize streamflow data. This package summarizes continuous daily mean streamflow data into various daily, monthly, annual, and long-term statistics, completes trending and frequency analyses.

This vignette explores how to obtain and integrate U.S. Geological Survey (USGS) streamflow data (available through the [National Water Information System (NWIS)](https://waterdata.usgs.gov/nwis)) using the [`dataRetrieval` R package](https://cran.r-project.org/package=dataRetrieval) with the various `fasstr` functions.

This vignette will use the following packages:

```{r}
library(dataRetrieval)
library(tidyhydat)
library(fasstr)
library(dplyr)
```

## dataRetrieval

`dataRetrieval` an R package with a collection of functions to retrieve hydrologic and water quality data hosted on USGS and U.S. Environmental Protection Agency (EPA) web services. USGS hydrologic data are pulled from <https://waterservices.usgs.gov/> and <https://waterdata.usgs.gov/nwis>. Water quality data are obtained from the Water Quality Portal <https://www.waterqualitydata.us/>. More information on using `dataRetrieval` can be found in the [vignette](https://cran.r-project.org/web/packages/dataRetrieval/vignettes/dataRetrieval.html).

This package is another useful tool for hydrologists, especially for those looking to use USGS streamflow hydrometric station data with `fasstr` or other streamflow packages and analyses. You can search for stations on the NWIS website <https://waterdata.usgs.gov/nwis/rt>. Unlike the `tidyhydat` package to obtain  Environment and Climate Change Canada's (ECCC) HYDAT historical discharge data, `dataRetrieval` pulls all data from the NWIS web services and does not require downloading and storing a database.

As USGS stations collect more than just streamflow, a parameter code is required for many of the functions to choose the appropriate data type. The parameter code for daily mean discharge data is "00060". The units of this discharge parameter are in US customary cubic feet per second (cfs). While using US customary units is not an issue for many `fasstr` functions, as some plots titles or column headers can simply be renamed, there are some functions that do metric unit conversions so some adaptations will be required (see sections below).

Here is a quick workflow demo showing how to get hydrometric station information, and daily and annual peak data using `dataRetrieval`, adapted from the [vignette](https://cran.r-project.org/web/packages/dataRetrieval/vignettes/dataRetrieval.html):

```{r, eval=FALSE}
# Hoko River Near Sekiu, WA (12043300)
site_number <- "12043300" 

# Get site information (ie. name, coordinates, drainage basin area, etc.)
site_info <- readNWISsite(site_number)

# Get daily discharge data
# Requires parameter code for discharge (00060), in cubic feet per second (cfs)
usgs_data <- readNWISdv(siteNumbers = site_number,
                        parameterCd = "00060", 
                        startDate = "1980-01-01",
                        endDate = "2019-12-31")

# Get annual instantaneous peak discharge data (based on water years (Oct-Sep))
peak_data <- readNWISpeak(siteNumbers = site_number)

```


## Converting to NWIS data metric and tidyhydat/fasstr format

For metric discharge analyses using NWIS data from the USGS, discharge values can easily be converted to cms from cfs (0.028317 cms per cfs) for use in fasstr.

The default dataRetrieval format is as follows:

```{r, eval=TRUE}
usgs_data <- readNWISdv(siteNumbers = "12043300", parameterCd = "00060")
head(usgs_data)
```

#### Keeping NWIS column names

To keep the NWIS data column names for fasstr functions, just the convert the discharge units and send to fasstr functions using the 'values' argument. Date column names are the same in both NWIS and HYDAT data. Using base R:

```{r, eval=FALSE}
# Get cfs data
usgs_data <- readNWISdv(siteNumbers = "12043300",
                        parameterCd = "00060")

# Convert cfs to cms
usgs_data$X_00060_00003 <- usgs_data$X_00060_00003 * 0.028317

# Set `values` and 'groups' (if multiple stations) to the appropriate column names
plot_flow_data(data = usgs_data,
               values = X_00060_00003,
               groups = site_no)
```

and using dplyr and pipelines:

```{r, eval=FALSE}
# One station
usgs_data <- readNWISdv(siteNumbers = "12043300", 
                        parameterCd = "00060") %>% 
  mutate(X_00060_00003 = X_00060_00003 * 0.028317)
plot_flow_data(data = usgs_data,
               values = X_00060_00003)

# Multiple stations
usgs_data <- readNWISdv(siteNumbers = c("12043300", "12048000"),
                        parameterCd = "00060") %>% 
  mutate(X_00060_00003 = X_00060_00003 * 0.028317)
plot_flow_data(data = usgs_data,
               values = X_00060_00003,
               groups = site_no)
```

#### Renaming column names

NWIS column names can also be changed to match those with HYDAT data formatting (ie. 'STATION_NUMBER', 'Date', and 'Value') to work easily with fasstr. Renaming 'site_no' to 'STATION_NUMBER' is not necessary if one station is used, but is useful if data contains multiple stations. Using base R:

```{r, eval=FALSE}
# Get the data
usgs_data <- readNWISdv(siteNumbers = "12043300", parameterCd = "00060")

# Rename columns
names(usgs_data)[names(usgs_data) == 'site_no'] <- 'STATION_NUMBER'
names(usgs_data)[names(usgs_data) == 'X_00060_00003'] <- 'Value'

# Convert cfs to cms
usgs_data$Value <- usgs_data$Value * 0.028317

# Use a fasstr function
calc_annual_peaks(usgs_data)
```

and using dplyr and pipelines:

```{r, eval=FALSE}
# One station
usgs_data <- readNWISdv(siteNumbers = "12043300", 
                        parameterCd = "00060") %>% 
  rename(Value = X_00060_00003) %>% 
  mutate(Value = Value * 0.028317)

# Multiple stations
usgs_data <- readNWISdv(siteNumbers = c("12043300", "12048000"), 
                        parameterCd = "00060") %>% 
  rename(STATION_NUMBER = site_no, 
         Value = X_00060_00003) %>% 
  mutate(Value = Value * 0.028317)

# Use a fasstr function
calc_annual_peaks(usgs_data)
```


#### Converting drainage basin areas

Some functions require a drainage basin area to calculate discharge as area-based runoff yield statistics (in depth units). The NWIS upstream drainage basin area can be found in the site information using the dataRetrieval `readNWISsite()` function and converted from square miles to square kilometres (2.58999 sqkm per sqmile):

```{r, eval=FALSE}
# Get station information
site_info <- readNWISsite("12043300")

# Get the drainage area and convert to sq. km
basin_area_NWIS <- site_info$drain_area_va * 2.58999

# Get data and apply to fasstr function
readNWISdv(siteNumbers = "12043300", 
           parameterCd = "00060") %>% 
  rename(Value = X_00060_00003) %>% 
  mutate(Value = Value * 0.028317) %>% 
  plot_daily_cumulative_stats(basin_area = basin_area_NWIS, 
                              use_yield = TRUE)
```

Here is one way of providing multiple basin areas for multiple stations:

```{r, eval=FALSE}
# Get station information
site_info <- readNWISsite(c("12043300", "12048000")) %>% 
  select(site_no, drain_area_va) %>% 
  mutate(Basin_Area_sqkm = drain_area_va * 2.58999)

# Create named vector of areas to send to basin area
areas <- setNames(site_info$Basin_Area_sqkm, site_info$site_no)

# Get data and apply to fasstr function
readNWISdv(siteNumbers = c("12043300", "12048000"), 
           parameterCd = "00060") %>% 
  rename(Value = X_00060_00003, 
         STATION_NUMBER = site_no) %>% 
  mutate(Value = Value * 0.028317) %>% 
  plot_daily_cumulative_stats(use_yield = TRUE,
                              basin_area = areas)
```

#### Handy formatting functions

Functions can also be written if conversions will frequently occur within workflows. The first example converts data already downloaded using dataRetrieval's `readNWISdv()` function:

```{r, eval=FALSE}
# Create function to convert to metric and HYDAT structure
usgs_to_hydat <- function(data){
  names(data)[names(data) == 'site_no'] <- 'STATION_NUMBER'
  names(data)[names(data) == 'X_00060_00003'] <- 'Value'
  data$Value <- data$Value * 0.028317
  return(data)
}

# Get data, convert, and send to fasstr function:
# in base R
usgs_data <- readNWISdv("12043300")
usgs_data <- usgs_to_hydat(usgs_data)
calc_annual_stats(usgs_data)

# and in a pipeline
readNWISdv("12043300") %>% 
  usgs_to_hydat() %>% 
  calc_annual_stats()
```

This second function integrates `readNWISdv()` with the first function that will both download and convert the data for fasstr functions by supplying NWIS site numbers:

```{r, eval=FALSE}
# Create function to download NWIS data and convert to metric and HYDAT structure
readNWISdv_hydat <- function(siteNumbers){
  data <- readNWISdv(siteNumbers = siteNumbers, parameterCd = "00060")
  names(data)[names(data) == 'site_no'] <- 'STATION_NUMBER'
  names(data)[names(data) == 'X_00060_00003'] <- 'Value'
  data$Value <- data$Value * 0.028317
  return(data)
}

# Get data, convert, and send to fasstr function:
# in base R
usgs_data <- readNWISdv_hydat("12043300")
calc_annual_stats(usgs_data)

# and in a pipeline
readNWISdv_hydat("12043300") %>% 
  calc_annual_stats()
```

This third function is the same as the second, but adds a column of drainage basin areas, in sq km, extracted from NWIS:

```{r, eval=FALSE}
# Create function to download NWIS data and convert to metric and HYDAT structure
readNWISdv_hydat <- function(siteNumbers){
  data <- readNWISdv(siteNumbers = siteNumbers, parameterCd = "00060")
  names(data)[names(data) == 'site_no'] <- 'STATION_NUMBER'
  names(data)[names(data) == 'X_00060_00003'] <- 'Value'
  data$Value <- data$Value * 0.028317
  
  basin_area <- readNWISsite(siteNumbers = siteNumbers)
  basin_area$Basin_Area_sqkm <- basin_area$drain_area_va * 2.58999
  basin_area <- data.frame(STATION_NUMBER = basin_area$site_no,
                           Basin_Area_sqkm = basin_area$Basin_Area_sqkm)
  
  data <- merge(data, basin_area, by = "STATION_NUMBER")
  
  return(data)
}

# Get data and convert data:
usgs_data <- readNWISdv_hydat("12043300")
```


## Working in US customary units

in NWIS data from the USGS and water years (water year designated by the calendar year it ends in, ie. 2010 for water year 2009-2010)


#### Keeping NWIS column names

NWIS data column names can be used with fasstr functions, with using `values = X_00060_00003` or `values = 'X_00060_00003'`. If multiple stations `groups = site_no` or `groups = 'site_no'`. Date column names are the same in both NWIS and HYDAT data. 

```{r, eval=FALSE}
# One station
usgs_data <- readNWISdv(siteNumbers = "12043300", 
                        parameterCd = "00060")
plot_flow_data(data = usgs_data,
               values = X_00060_00003)

# Multiple stations
usgs_data <- readNWISdv(siteNumbers = c("12043300", "12048000"),
                        parameterCd = "00060")
plot_flow_data(data = usgs_data,
               values = X_00060_00003,
               groups = site_no)

# Calculate stats using water years
calc_longterm_mean(data = usgs_data, 
                   values = X_00060_00003,
                   water_year_start = 10, 
                   complete_years = TRUE)
```

#### Renaming NWIS column names

NWIS column names can also be changed to match those with HYDAT data formatting (ie. 'STATION_NUMBER', 'Date', and 'Value') to work easily with fasstr. Using original names is good for quick function use, but renaming to the HYDAT convention makes using multiple fasstr functions easier. Renaming 'site_no' to 'STATION_NUMBER' is not necessary if one station is used, but is useful if data contains multiple stations. Using base R:

```{r, eval=FALSE}
# Get the data
usgs_data <- readNWISdv(siteNumbers = "12043300", parameterCd = "00060")

# Rename columns
names(usgs_data)[names(usgs_data) == 'site_no'] <- 'STATION_NUMBER'
names(usgs_data)[names(usgs_data) == 'X_00060_00003'] <- 'Value'
```

and using dplyr and pipelines:

```{r, eval=FALSE}
# One station
usgs_data <- readNWISdv(siteNumbers = "12043300", 
                        parameterCd = "00060") %>% 
  rename(Value = X_00060_00003) %>% 
  mutate(Value = Value * 0.028317)

# Multiple stations
usgs_data <- readNWISdv(siteNumbers = c("12043300", "12048000"), 
                        parameterCd = "00060") %>% 
  rename(STATION_NUMBER = site_no, 
         Value = X_00060_00003) %>% 
  mutate(Value = Value * 0.028317)
```

#### Example workflows


Basic, no renaming or anything
```{r, eval=FALSE}
# Download and format NWIS data
usgs_data <- readNWISdv(siteNumbers = "12043300", 
                        parameterCd = "00060") %>% 
  rename(Value = X_00060_00003)

# Fill in some data gaps and add some columns
usgs_data <- usgs_data %>% 
  fill_missing_dates(water_year_start = 10) %>% 
  add_date_variables(water_year_start = 10) %>% 
  add_rolling_means(roll_days = 7)

# Send to various fasstr functions
screen_flow_data(usgs_data, 
                 water_year_start = 10)

calc_longterm_daily_stats(usgs_data, 
                          water_year_start = 10,
                          complete_years = TRUE)

calc_longterm_mean(usgs_data, 
                   water_year_start = 10, 
                   complete_years = TRUE)

calc_annual_lowflows(usgs_data, 
                     water_year_start = 10)
```

List all functions?

Some will require renaming of plot titles/values






#### Use of drainage basin areas




```{r, eval=FALSE}
readNWISdv(siteNumbers = "12043300", 
           parameterCd = "00060") %>% 
  rename(Value = X_00060_00003) %>% 
  fill_missing_dates(water_year_start = 10) %>% 
  add_date_variables(water_year_start = 10) %>% 
  add_seasons(water_year_start = 10,
              seasons_length = 3) %>% 
  add_rolling_means() %>% 
  add_daily_volume() %>% 
    rename(Volume_cf = Volume_m3) %>% 
  add_cumulative_volume(water_year_start = 10) %>% 
    rename(Cumul_Volume_cf = Cumul_Volume_m3) %>% 
  add_daily_yield(basin_area = 10.3 * 2323.2) %>% 
    rename(Yield_in = Yield_mm) %>%  
  add_cumulative_yield(water_year_start = 10, basin_area = 10.3 * 2323.2) %>% 
    rename(Cumul_Yield_in = Cumul_Yield_mm)
```




#### Examples

-   basic, no basin_area
-   Simple calcs and plots (no conversions), screen, plot, plot_means, daily, low flow frequency, flow duration, trending?, custom peak frequency

#### Frequency Examples

#### Basin Area Examples

-   area-based runoff/yield and cumulative examples
-   basin area tricks


- cumulative yield units, toooo large, acre feet? what units

USE WASHINGTON EXAMPLES - COAST AND INTERIOR

12048000 DUNGENESS RIVER NEAR SEQUIM, WA medium coastal, very mixed regime 12045500 ELWHA RIVER AT MCDONALD BR NEAR PORT ANGELES, WA large coastal, mixed 12447390 ANDREWS CREEK NEAR MAZAMA, WA small nival 12404500 KETTLE RIVER NEAR LAURIER, WA large nival 12442500 SIMILKAMEEN RIVER NEAR NIGHTHAWK, WA nival 12043300 HOKO RIVER NEAR SEKIU, WA small coastal, pluvial 12212050 FISHTRAP CREEK AT FRONT STREET AT LYNDEN, WA pluvial - Abbotsford
