---
title: "Using USGS dataRetrieval R package with fasstr"
output: 
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{Using USGS dataRetrieval R package with fasstr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r options, include=FALSE}
knitr::opts_chunk$set(eval = nzchar(Sys.getenv("hydat_eval")),
                      warning = FALSE, 
                      message = FALSE)
```

`fasstr`, the Flow Analysis Summary Statistics Tool for R, is a set of [R](https://www.r-project.org/) functions to tidy, summarize, analyze, trend, and visualize streamflow data. This package summarizes continuous daily mean streamflow data into various daily, monthly, annual, and long-term statistics, completes trending and frequency analyses.

This vignette explores how to obtain and integrate U.S. Geological Survey (USGS) streamflow data, available through the [National Water Information System (NWIS)](https://waterdata.usgs.gov/nwis) using the [`dataRetrieval` R package](https://cran.r-project.org/package=dataRetrieval), with the various `fasstr` functions.

This vignette will use the following packages:

```{r, eval=TRUE}
library(fasstr)
library(dataRetrieval) # for getting USGS NWIS data
library(tidyhydat) # for getting ECCC HYDAT data
library(dplyr) # for data wrangling and pipelines
library(ggplot2) # for modifying fasstr plots
```

## dataRetrieval Package

`dataRetrieval` is an R package with a collection of functions to retrieve hydrologic and water quality data hosted on USGS and U.S. Environmental Protection Agency (EPA) web services. USGS hydrologic data are pulled from <https://waterservices.usgs.gov/> and <https://waterdata.usgs.gov/nwis>. Water quality data are obtained from the Water Quality Portal <https://www.waterqualitydata.us/>. More information on using `dataRetrieval` can be found in the [vignette](https://cran.r-project.org/web/packages/dataRetrieval/vignettes/dataRetrieval.html).

This package is another useful tool for hydrologists, especially for those looking to use USGS streamflow hydrometric station data with `fasstr` or other streamflow packages and analyses. You can search for stations on the NWIS website <https://waterdata.usgs.gov/nwis/rt>. Unlike the `tidyhydat` package to obtain  Environment and Climate Change Canada's (ECCC) HYDAT historical discharge data, `dataRetrieval` pulls all data from the NWIS web services and does not require downloading and storing a database.

As USGS stations collect more than just streamflow, a parameter code is required for many of the functions to choose the appropriate data type. The parameter code for daily mean discharge data is "00060". The units of this discharge parameter are in US customary cubic feet per second (cfs). While using US customary units is not an issue for many `fasstr` functions, as units do not matter for many analyses or plots titles and column headers can simply be renamed, there are some functions that use internal metric unit conversions so some adaptations will be required (see sections below).

Here is a quick workflow demo showing how to get hydrometric station information, and daily and annual peak data using `dataRetrieval`, adapted from the [vignette](https://cran.r-project.org/web/packages/dataRetrieval/vignettes/dataRetrieval.html):

```{r, eval=FALSE}
# Hoko River Near Sekiu, WA (12043300)
site_number <- "12043300" 

# Get site information (ie. name, coordinates, drainage basin area, etc.)
site_info <- readNWISsite(site_number)

# Get daily discharge data
# Requires parameter code for discharge (00060), in cubic feet per second (cfs)
usgs_data <- readNWISdv(siteNumbers = site_number,
                        parameterCd = "00060", 
                        startDate = "1980-01-01",
                        endDate = "2019-12-31")

# Get annual instantaneous peak discharge data (based on water years (Oct-Sep))
peak_data <- readNWISpeak(siteNumbers = site_number)

```


## Converting to metric and fasstr default columns

For metric discharge analyses using NWIS data from the USGS, discharge values can easily be converted to cms from cfs (0.028317 cms per cfs) for use in fasstr.

The default dataRetrieval NWIS data column formats are as follows:

```{r, eval=TRUE}
usgs_data <- readNWISdv(siteNumbers = "12043300", parameterCd = "00060")
head(usgs_data)
```

### Keeping NWIS column names

To keep the NWIS data column names for fasstr functions, just remember to convert the discharge units and then send to fasstr functions using the 'values' argument. Date column names are the same in both NWIS and HYDAT data. Here is an example of this situation using base R:

```{r, eval=FALSE}
# Get cfs data
usgs_data <- readNWISdv(siteNumbers = "12043300",
                        parameterCd = "00060")

# Convert cfs to cms
usgs_data$X_00060_00003 <- usgs_data$X_00060_00003 * 0.028317

# Set `values` and 'groups' (if multiple stations) to the appropriate column names
plot_flow_data(data = usgs_data,
               values = X_00060_00003,
               groups = site_no)
```

and using dplyr and pipelines:

```{r, eval=FALSE}
# One station
usgs_data <- readNWISdv(siteNumbers = "12043300", 
                        parameterCd = "00060") %>% 
  mutate(X_00060_00003 = X_00060_00003 * 0.028317)
plot_flow_data(data = usgs_data,
               values = X_00060_00003)

# Multiple stations
usgs_data <- readNWISdv(siteNumbers = c("12043300", "12048000"),
                        parameterCd = "00060") %>% 
  mutate(X_00060_00003 = X_00060_00003 * 0.028317)
plot_flow_data(data = usgs_data,
               values = X_00060_00003,
               groups = site_no)
```

### Renaming NWIS column names

NWIS column names can also be changed to match those with tidyhydat and fasstr's defaults data column names (ie. 'STATION_NUMBER', 'Date', and 'Value') to work easily with fasstr. Renaming 'site_no' to 'STATION_NUMBER' is not necessary if one station is used, but is useful if data contains multiple stations. Here is an example of this situation using base R:

```{r, eval=FALSE}
# Get the data
usgs_data <- readNWISdv(siteNumbers = "12043300", parameterCd = "00060")

# Rename columns
names(usgs_data)[names(usgs_data) == 'site_no'] <- 'STATION_NUMBER'
names(usgs_data)[names(usgs_data) == 'X_00060_00003'] <- 'Value'

# Convert cfs to cms
usgs_data$Value <- usgs_data$Value * 0.028317

# Use a fasstr function
calc_annual_peaks(usgs_data)
```

and using dplyr and pipelines:

```{r, eval=FALSE}
# One station
usgs_data <- readNWISdv(siteNumbers = "12043300", 
                        parameterCd = "00060") %>% 
  rename(Value = X_00060_00003) %>% 
  mutate(Value = Value * 0.028317)

# Multiple stations
usgs_data <- readNWISdv(siteNumbers = c("12043300", "12048000"), 
                        parameterCd = "00060") %>% 
  rename(STATION_NUMBER = site_no, 
         Value = X_00060_00003) %>% 
  mutate(Value = Value * 0.028317)

# Use a fasstr function
calc_annual_peaks(usgs_data)
```


### Getting and converting drainage basin areas

Some fasstr functions require a drainage basin area to calculate discharge as area-based runoff yield statistics (in depth units, commonly millimetres in Canada). The NWIS upstream drainage basin area can be found in the site information using the dataRetrieval `readNWISsite()` function and then converted from square miles to square kilometres (2.58999 sqkm per sqmile):

```{r, eval=FALSE}
# Get station information
site_info <- readNWISsite("12043300")

# Get the drainage area and convert to sq. km
basin_area_NWIS <- site_info$drain_area_va * 2.58999

# Get data and apply to fasstr function
readNWISdv(siteNumbers = "12043300", 
           parameterCd = "00060") %>% 
  rename(Value = X_00060_00003) %>% 
  mutate(Value = Value * 0.028317) %>% 
  plot_daily_cumulative_stats(basin_area = basin_area_NWIS, 
                              use_yield = TRUE)
```

Here is one way of providing multiple basin areas for multiple stations:

```{r, eval=FALSE}
# Get station information
site_info <- readNWISsite(c("12043300", "12048000")) %>% 
  select(site_no, drain_area_va) %>% 
  mutate(Basin_Area_sqkm = drain_area_va * 2.58999)

# Create named vector of areas to send to basin area
areas <- setNames(site_info$Basin_Area_sqkm, site_info$site_no)

# Get data and apply to fasstr function
readNWISdv(siteNumbers = c("12043300", "12048000"), 
           parameterCd = "00060") %>% 
  rename(Value = X_00060_00003, 
         STATION_NUMBER = site_no) %>% 
  mutate(Value = Value * 0.028317) %>% 
  plot_daily_cumulative_stats(use_yield = TRUE,
                              basin_area = areas)
```

### Handy formatting functions

Functions can also be created if conversions will frequently occur within workflows. The first example here converts and formats data already downloaded using dataRetrieval's `readNWISdv()` function for fasstr:

```{r, eval=FALSE}
# Create function to convert to metric and HYDAT structure
usgs_to_hydat <- function(data){
  names(data)[names(data) == 'site_no'] <- 'STATION_NUMBER'
  names(data)[names(data) == 'X_00060_00003'] <- 'Value'
  data$Value <- data$Value * 0.028317
  return(data)
}

# Get data, convert, and send to fasstr function:
# in base R
usgs_data <- readNWISdv("12043300")
usgs_data <- usgs_to_hydat(usgs_data)
calc_annual_stats(usgs_data)

# and in a pipeline
readNWISdv("12043300") %>% 
  usgs_to_hydat() %>% 
  calc_annual_stats()
```

This second function integrates `readNWISdv()` with the first function that will both download and convert the data for fasstr functions by supplying NWIS site numbers:

```{r, eval=FALSE}
# Create function to download NWIS data and convert to metric and HYDAT structure
readNWISdv_hydat <- function(siteNumbers){
  data <- readNWISdv(siteNumbers = siteNumbers, parameterCd = "00060")
  names(data)[names(data) == 'site_no'] <- 'STATION_NUMBER'
  names(data)[names(data) == 'X_00060_00003'] <- 'Value'
  data$Value <- data$Value * 0.028317
  return(data)
}

# Get data, convert, and send to fasstr function:
# in base R
usgs_data <- readNWISdv_hydat("12043300")
calc_annual_stats(usgs_data)

# and in a pipeline
readNWISdv_hydat("12043300") %>% 
  calc_annual_stats()
```

This third function is the same as the second, but adds a column of drainage basin areas, in sq km, extracted from NWIS:

```{r, eval=FALSE}
# Create function to download NWIS data and convert to metric and HYDAT structure
readNWISdv_hydat <- function(siteNumbers){
  data <- readNWISdv(siteNumbers = siteNumbers, parameterCd = "00060")
  names(data)[names(data) == 'site_no'] <- 'STATION_NUMBER'
  names(data)[names(data) == 'X_00060_00003'] <- 'Value'
  data$Value <- data$Value * 0.028317
  
  basin_area <- readNWISsite(siteNumbers = siteNumbers)
  basin_area$Basin_Area_sqkm <- basin_area$drain_area_va * 2.58999
  basin_area <- data.frame(STATION_NUMBER = basin_area$site_no,
                           Basin_Area_sqkm = basin_area$Basin_Area_sqkm)
  
  data <- merge(data, basin_area, by = "STATION_NUMBER")
  
  return(data)
}

# Get data and convert data:
usgs_data <- readNWISdv_hydat("12043300")
```


## Working in US customary units

units customary provided by NWIS - cfs and sq miles

CFS will work with most functions without a problem. most just take data provided and provide means, percentiles, etc. For tables, can assume if there are no units in column names that units are from supplied data (cfs). Axes labeled 'Discharge (cms)' on plots can simply be renamed by extracting the plot object from its listed object (using double square brackets, eg. `[[1]]`, or `magrittr::extract2(1)` in a pipeline) and using ggplot2 functions (eg. `+ ylab('Discharge, cubic feet per second')`). These column names and plot axes can easily be changed to match the units (see below for examples). If units are in Volume_mm it can be interpeteated as cubic feet volume and renamed as such. Only if column names or plots titles are in Yield will there be additonal conversions required to get the appropriate inches. See the section below on drainage basin area conversions.

in NWIS data from the USGS and water years (water year designated by the calendar year it ends in, ie. 2010 for water year 2009-2010) `water_year_start = 10`

### Keeping NWIS column names

NWIS data column names can easily be used with fasstr functions, with setting `values = X_00060_00003` or `values = 'X_00060_00003'` in fasstr functions to identify the flow values column. If there are multiple stations in a data frame, the stations can be grouped separately by identifying them using `groups = site_no` or `groups = 'site_no'` in fasstr functions. Date column names are the same in both NWIS and HYDAT data. 

```{r, eval=FALSE}
# One station
usgs_data <- readNWISdv(siteNumbers = "12043300", 
                        parameterCd = "00060")
plot_flow_data(data = usgs_data,
               values = X_00060_00003)

# Multiple stations
usgs_data <- readNWISdv(siteNumbers = c("12043300", "12048000"),
                        parameterCd = "00060")
plot_flow_data(data = usgs_data,
               values = X_00060_00003,
               groups = site_no)

# Calculate stats using water years
calc_longterm_mean(data = usgs_data, 
                   values = X_00060_00003,
                   water_year_start = 10, 
                   complete_years = TRUE)
```

### Renaming NWIS column names

NWIS column names can also be changed to match those with tidyhydat and fasstr's defaults data column names (ie. 'STATION_NUMBER', 'Date', and 'Value') to work easily with fasstr. Using original names is good for quick function use, but renaming to the HYDAT convention makes using multiple fasstr functions easier. Renaming 'site_no' to 'STATION_NUMBER' is not necessary if one station is used, but is useful if data contains multiple stations. Here is an example of this situation using base R:

```{r, eval=FALSE}
# Get the data
usgs_data <- readNWISdv(siteNumbers = "12043300", parameterCd = "00060")

# Rename columns
names(usgs_data)[names(usgs_data) == 'site_no'] <- 'STATION_NUMBER'
names(usgs_data)[names(usgs_data) == 'X_00060_00003'] <- 'Value'
```

and using dplyr and pipelines:

```{r, eval=FALSE}
# One station
usgs_data <- readNWISdv(siteNumbers = "12043300", 
                        parameterCd = "00060") %>% 
  rename(Value = X_00060_00003)

# Multiple stations
usgs_data <- readNWISdv(siteNumbers = c("12043300", "12048000"), 
                        parameterCd = "00060") %>% 
  rename(STATION_NUMBER = site_no, 
         Value = X_00060_00003)
```


### Use of drainage basin areas

Some fasstr functions use require drainage basin areas to calculate discharge as area-based runoff yield statistics (in depth units). Since fasstr internally converts discharge using metric conversions, this makes conversions of US customary cubic feet per second and square miles to inches depth not as simple. Luckily, if your discharge units are in cubic feet per second and basin ares in square miles, a simple workaround can be applied to get inches depth: multiply the basin area by 2323.2. This number is derived from three unit conversions that magically provide the proper conversion:

```{r}
# sqkm/sqmiles / (inches/mm * cms/cfs)
2.58999 / (0.0393701 * 0.028316846592)
```

So if using any functions that result in yield values (typically 'yield' or 'cumulative' functions that with columns or axes with yield names when using `use_yield = TRUE`), then the multiplier needs to be applied to square mile basin areas. NWIS upstream drainage basin areas can be found in the site information using the dataRetrieval `readNWISsite()` function. The following is an example of how to use the workaround, with appropriate renaming of columns or plot axes:

```{r, eval=FALSE}
# Get a NWIS basin area
area_sqmile <- dataRetrieval::readNWISsite(siteNumbers = "12043300") %>% 
  pull(drain_area_va)

# Get NWIS data
usgs_data <- readNWISdv(siteNumbers = "12043300", 
                        parameterCd = "00060") %>% 
  rename(Value = X_00060_00003)

# Add daily yield in inches, with 2323.2 basin area multiplier
add_daily_yield(data = usgs_data,
                basin_area = area_sqmile * 2323.2) %>% 
  rename(Yield_inch = Yield_mm)

# Calculate total annual inches, with multiplier
calc_annual_cumulative_stats(data = usgs_data, 
                             use_yield = TRUE,
                             basin_area = area_sqmile * 2323.2) %>% 
  rename(Total_Yield_inch = Total_Yield_mm)

# Plot daily cumulative discharge, with multiplier
plot_daily_cumulative_stats(data = usgs_data, 
                            use_yield = TRUE,
                            basin_area = area_sqmile * 2323.2) %>% 
  magrittr::extract2(1) + # extracts plot from list
  ggplot2::ylab("Runoff, inch")
```

### Examples: data wrangling

For the rest of the vignette examples, we will set up data frames of daily stream discharge from dataRetrieval, called `usgs_data` and `usgs_data_multiple`, and rename the columns to match the fasstr defaults (i.e. STATION_NUMBER, Date, and Value). We will also extract a basin area for the `usgs_data` data set.

```{r, eval=FALSE}
# Get daily data for examples
usgs_data <- readNWISdv(siteNumbers = "12043300", 
                        parameterCd = "00060") %>% 
  rename(Value = X_00060_00003,
         STATION_NUMBER = site_no)

# Get drainage basin area for examples
usgs_basin_area <- dataRetrieval::readNWISsite(siteNumbers = "12043300") %>% 
  pull(drain_area_va)

# Get daily data for examples with multiple stations
usgs_data_multiple <- readNWISdv(siteNumbers = c("12043300", "12048000"), 
                                 parameterCd = "00060") %>% 
  rename(Value = X_00060_00003,
         STATION_NUMBER = site_no)
```

To be consistent with many USGS streamflow analyses, these examples will use Water Years (from Oct-Sep) instead of Calendar Years by setting `water_year_start = 10` in the necessary functions. With fasstr, a water year is identified by the calendar year in which it ends. So for example, a water year from October 2019 to September 2020 will be identified as 2020.

### Examples: tidying functions

There are several functions that are used to prepare your flow data set for your own analysis by adding columns of various date variables, adding rolling means, and adding cumulative discharges, amongst others. These functions begin with `add_` or `fill_` and add columns or rows, respectively, to your data frame of daily flow values. See the documentation for more info on the functions. The following example shows how to apply the tidying functions in a continuous pipeline and using water years. Note how the basin area correction is applied in the two `_cumulative_` functions and columns are renamed to match the appropriate units where necessary.

```{r, eval=FALSE}
usgs_data %>% 
  fill_missing_dates(water_year_start = 10) %>% 
  add_date_variables(water_year_start = 10) %>% 
  add_seasons(water_year_start = 10,
              seasons_length = 3) %>% 
  add_rolling_means() %>% 
  add_daily_volume() %>% 
  rename(Volume_cuft = Volume_m3) %>% 
  add_cumulative_volume(water_year_start = 10) %>% 
  rename(Cumul_Volume_cuft = Cumul_Volume_m3) %>% 
  add_daily_yield(basin_area = usgs_basin_area * 2323.2) %>% 
  rename(Yield_inch = Yield_mm) %>%  
  add_cumulative_yield(water_year_start = 10, 
                       basin_area = usgs_basin_area * 2323.2) %>% 
  rename(Cumul_Yield_inch = Cumul_Yield_mm)
```

### Examples: flow data screening

It may be useful to explore the data quality and availability before using the data for analysis or for interest. Some of the screening functions can help you screen for outliers and missing data. Note the renaming of axis labels in the plot examples to the necessary units with one or multiple station plots.

```{r, eval=FALSE}
# Get the number of missing dates and some basic summary statistics (calculated regardless 
# of missing dates)
screen_flow_data(usgs_data, 
                 water_year_start = 10)

# Plot the missing dates
plot_missing_dates(usgs_data, 
                   water_year_start = 10)

# Plot the data, extract the plot, and rename the axis
plot_flow_data(usgs_data)[[1]]+
  ylab("Discharge, cfs")

# Alternatively, plot the data, and rename the axis label by modifying the ggplot code
usgs_daily_plot <- plot_flow_data(usgs_data)
usgs_daily_plot$Daily_Flows$labels$y <- "Discharge, cfs"
usgs_daily_plot

######################
##################
plot_flow_data(usgs_data_multiple, 
               one_plot = TRUE)[[1]]+
  ylab("Discharge, cfs")

usgs_daily_multiple <- plot_flow_data(usgs_data_multiple)
usgs_daily_multiple$`12043300_Daily_Flows`$labels$y <- "Discharge, cfs"
usgs_daily_multiple$`12048000_Daily_Flows`$labels$y <- "Discharge, cfs"
usgs_daily_multiple
```

### Examples: summary statistics tables and plots

The majority of the `fasstr` functions produce statistics over a certain time period, either long-term, annually, monthly, or daily. These statistics are produced using the `calc_*` functions and can be visualized using their corresponding `plot_*` functions. Most of the `calc_` functions do not require any changes to column names, unless they require as drainage basin area, where the multiplier fix will be applied. Most of the plots will require renaming axes as noted.


```{r, eval=FALSE}
# Calculate long-term and long-term monthly summary statistics based on daily data
# using only years with complete data
calc_longterm_daily_stats(usgs_data, 
                          water_year_start = 10,
                          complete_years = TRUE)

# Calculate the long-term mean annual discharge (MAD) using only years with complete 
# data and calculate 5%, 10%, and 20% MADs
calc_longterm_mean(usgs_data, 
                   water_year_start = 10, 
                   complete_years = TRUE,
                   percent_MAD = c(5,10,20))

# Plot flow duration curves for each month and combining summer months, using only 
# complete years, and renaming the axis title
plot_flow_duration(usgs_data, 
                   water_year_start = 10, 
                   complete_years = TRUE,
                   custom_months = 7:9,
                   custom_months_label = "Summer")[[1]]+
  ylab("Discharge, cubic feet per second")

# Plot annual summary statistics, including annual 10th and 90th percentiles, starting
# in 1996 and making the axis logarithmic, and renaming the axis title
plot_annual_stats(usgs_data, 
                  water_year_start = 10,
                  percentiles = c(10,90),
                  start_year = 1996,
                  log_discharge = TRUE)[[1]]+
  ylab("Discharge, cubic feet per second")

# Calculate annual 7- and 30-day low flows
calc_annual_lowflows(usgs_data, 
                     roll_days = c(7,30),
                     water_year_start = 10)

# Plot the timing of half flows
plot_annual_flow_timing(usgs_data,
                        percent_total = 50,
                        water_year_start = 10, mon)




# calc monthly

# calc daiyl

# plot daily

# plot daily cumulative

# plot cumualtive in cubic feet to acre feet conversion? changing the ylabels

```


### Examples: frequency analyses

annual frequencies

```{r, eval=FALSE}
# Compute annual low flows frequencies with default arguments
lowflow_frequency <- compute_annual_frequencies(usgs_data,
                                                water_year_start = 10)
# Change the y-axis of the plot to US units (changing the object itself)
lowflow_frequency$Freq_Plot$labels$y <- "Discharge, cubic feet per second"

# Compute annual daily peak flows frequencies with default arguments
highflow_frequency <- compute_annual_frequencies(usgs_data,
                                                 water_year_start = 10,
                                                 use_max = TRUE,
                                                 roll_days = 1)
# Change the y-axis of the plot to US units (changing the object itself)
highflow_frequency$Freq_Plot$labels$y <- "Discharge, cubic feet per second"
highflow_frequency$Freq_Plot




# Compute summer low flow frequencies with default arguments
lowflow_7day <- compute_annual_frequencies(usgs_data,
                                           water_year_start = 10,
                                           months = 7:9,
                                           roll_days = 7)
# Extract the plot and change the y-axis title to US units (doesn't change the original)
lowflow_plot <- lowflow_7day$Freq_Plot +
  ylab("Discharge, cubic feet per second")
```

freq_quantile

custom frequencies

```{r, eval=FALSE}
# dataRetrieval function to get annual instantaneous peak data, grouped by water years
usgs_peaks <- readNWISpeak("12043300") %>% 
  mutate(Measure = "Inst. Peaks")

# Use the default column names for events and values for the custom analysis 
# and default arguments (log-PIII etc.)
flood_frequency <- compute_frequency_analysis(data = usgs_peaks,
                                              events = peak_dt,
                                              values = peak_va,
                                              measures = Measure,
                                              use_max = TRUE)

# Can also do the analysis with renaming parameters to match fasstr function, within a pipeline
flood_frequency <- readNWISpeak("12043300") %>% 
  select(Year = peak_dt, Value = peak_va) %>% 
  mutate(Measure = "Inst. Peaks") %>% 
  compute_frequency_analysis(use_max = TRUE)
```

### Examples: annual trending

trending example applyig the basin area correction, note of units

```{r, eval=FALSE}
trends <- compute_annual_trends(usgs_data, 
                                zyp_method = "zhang", 
                                zyp_alpha = 0.05,
                                basin_area = usgs_basin_area * 2323.2,
                                water_year_start = 10)
```

However, while values will be accurate, labels will not be for some. If want to use results, then here are some ways to convert the labels:

```{r, eval=FALSE}
# Rename plot names
names(trends) <- gsub('_m3', '_cuft', names(trends))
names(trends) <- gsub('_mm', '_inch', names(trends))

# Rename the Statistics in Data and Results
trends$Annual_Trends_Data$Statistic <- gsub('_m3', '_cuft', trends$Annual_Trends_Data$Statistic)
trends$Annual_Trends_Data$Statistic <- gsub('_mm', '_inch', trends$Annual_Trends_Data$Statistic)
trends$Annual_Trends_Results$Statistic <- gsub('_m3', '_cuft', trends$Annual_Trends_Results$Statistic)
trends$Annual_Trends_Results$Statistic <- gsub('_mm', '_inch', trends$Annual_Trends_Results$Statistic)

# Rename y-axes and titles in plots
for (i in names(trends[3:length(trends)])){
  if (all(trends[[i]]$labels$y == "Discharge (cms)")){
    trends[[i]]$labels$y <- gsub('cms', 'cfs', trends[[i]]$labels$y)
  }
  if (all(trends[[i]]$labels$y == "Yield (mm)")){
    trends[[i]]$labels$y <- gsub('mm', 'inch', trends[[i]]$labels$y)
    trends[[i]]$labels$title <- gsub('_mm', '_inch', trends[[i]]$labels$title)
  }
  if (all(trends[[i]]$labels$y == "Volume (cubic metres)")){
    trends[[i]]$labels$y <- gsub('metres', 'feet', trends[[i]]$labels$y)
    trends[[i]]$labels$title <- gsub('_m3', '_cuft', trends[[i]]$labels$title)
  }
}
```

A function can also be written if relabeling will frequently occur within workflows. With this example, you apply the results of `compute_annual_trends()` directly to this wrapper function and it will rename all appropriate labels and titles within the results.

```{r, eval=FALSE}
fasstr_trends_to_US_units <- function(trends){
  
  # Rename plot names
  names(trends) <- gsub('_m3', '_cuft', names(trends))
  names(trends) <- gsub('_mm', '_inch', names(trends))
  
  # Rename the Statistics in Data and Results
  trends$Annual_Trends_Data$Statistic <- gsub('_m3', '_cuft', trends$Annual_Trends_Data$Statistic)
  trends$Annual_Trends_Data$Statistic <- gsub('_mm', '_inch', trends$Annual_Trends_Data$Statistic)
  trends$Annual_Trends_Results$Statistic <- gsub('_m3', '_cuft', trends$Annual_Trends_Results$Statistic)
  trends$Annual_Trends_Results$Statistic <- gsub('_mm', '_inch', trends$Annual_Trends_Results$Statistic)
  
  # Rename y-axes and titles in plots
  for (i in names(trends[3:length(trends)])){
    if (all(trends[[i]]$labels$y == "Discharge (cms)")){
      trends[[i]]$labels$y <- gsub('cms', 'cfs', trends[[i]]$labels$y)
    }
    if (all(trends[[i]]$labels$y == "Yield (mm)")){
      trends[[i]]$labels$y <- gsub('mm', 'inch', trends[[i]]$labels$y)
      trends[[i]]$labels$title <- gsub('_mm', '_inch', trends[[i]]$labels$title)
    }
    if (all(trends[[i]]$labels$y == "Volume (cubic metres)")){
      trends[[i]]$labels$y <- gsub('metres', 'feet', trends[[i]]$labels$y)
      trends[[i]]$labels$title <- gsub('_m3', '_cuft', trends[[i]]$labels$title)
    }
  }
  
  return(trends)
}

# Apply the wrapper function
trends <- fasstr_trends_to_US_units(trends)

```


## To Do

Ask usgs 

- is acre-feet used often? any demos needed, especially for total volumes? cumulative yield units, toooo large, acre feet? what units
- other analyses?
- review vignette

USE WASHINGTON EXAMPLES - COAST AND INTERIOR

12048000 DUNGENESS RIVER NEAR SEQUIM, WA medium coastal, very mixed regime 12045500 ELWHA RIVER AT MCDONALD BR NEAR PORT ANGELES, WA large coastal, mixed 12447390 ANDREWS CREEK NEAR MAZAMA, WA small nival 12404500 KETTLE RIVER NEAR LAURIER, WA large nival 12442500 SIMILKAMEEN RIVER NEAR NIGHTHAWK, WA nival 12043300 HOKO RIVER NEAR SEKIU, WA small coastal, pluvial 12212050 FISHTRAP CREEK AT FRONT STREET AT LYNDEN, WA pluvial - Abbotsford
