---
title: "Fun with fasstr"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Fun with fasstr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
  
  
```{r options, include=FALSE}
knitr::opts_chunk$set(eval = nzchar(Sys.getenv("hydat_eval")),
                      warning = FALSE, 
                      message = FALSE)
```


Topics

* Tidyverse fun (daily with reference lines)
* Integration with other packages
* Groundwater Data
* Fitting data to other distributions (fasstr uses those in HEC-SSP, but others)


These examples use the following packages
```{r}
library(fasstr)
library(dplyr)
```

### Plotting with yields comparison
### Stats and plotting GW Levels
### adding lines to daily plots
### add realtime data to daily plot
### writing results with fasstr function within (quick excel file for exploration)
### Other probability distribution fitting with fasstr results
##### calculate low flows or high flows
##### test for independence, stationarity, homegeniety
##### frequency curve for plotting
##### fit distribution
https://rpubs.com/blakeobeans/fitdistrplus




fasstr data doesnt require objects to be certain classes, just requires a data frame with the time series (Dates and Values). 

These examples use the following packages
```{r}
library(fasstr)
library(dplyr)
```

### tidyhydat

tidyhydat is integrated into fasstr, but can use apart and bring into fasstr functions.


### USGS Data



### FlowScreen Timeseries

```{r}
data <- tidyhydat::hy_daily_flows(station_number = "08NM116") %>% 
  rename(SYM = Symbol,
         Flow = Value,
         ID = STATION_NUMBER) %>% 
  mutate(Agency = "WSC")
data <- FlowScreen::create.ts(data)
```


### lfstat objects

```{r}

```

